---
title: "Individual2-Lab4"
author: "Ritika Agrawal"
date: "March 20, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#File Read In
library(readxl)
library(tidyr)
library(dplyr)
library(plotly)
defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
str(defense)
```


```{r}
#Part One

#i
defClean <- defense %>%
  mutate(Name = as.factor(Name),Opponent_Opponent=as.factor(Opponent_Opponent)) %>%
  mutate_at(c("Tackles_Solo","Tackles_ASST","Tackles_TFL","Tackles_Sack","Turnover_FF","Turnover_FR","Turnover_INT","Pass_QBH","Pass_PB"),as.numeric)
  


nrow(defClean)
ncol(defClean)
head(defClean)

offClean <- offense %>%
   
  mutate(Name = as.factor(Name),Opponent_Opponent=as.factor(Opponent_Opponent)) %>%
  replace_na(list(.=0)) %>%
  mutate_at(c("Rushing_ATT","Rushing_YDS","Rushing_TD","Receiving_REC","Receiving_YDS","Receiving_TD","Passing_CMP-ATT","Passing_YDS","Passing_TD","Passing_INT"),as.numeric)

nrow(offClean)
ncol(offClean)

bioClean <- bio %>% 
  separate(col=Height,into=c("Feet","Inches"),sep="-") %>%
  mutate(Feet=as.numeric(Feet),Inches=as.numeric(Inches)) %>%
  mutate(Height_Inch = Feet*12+Inches) %>%
  select(-Feet,-Inches)

nrow(bioClean)
ncol(bioClean)

```

## Part 2
```{r}
defClean2 = pivot_longer(
  defClean,
  cols = Tackles_Solo:Pass_PB,
  names_to = "Statistic",
  values_to = "Count",
  values_drop_na = TRUE
)
head(defClean2)



```

Response: A new column, 'Statistic' has been added. 


```{r}
#qplot(defClean2$Count, geom="histogram") + facet_wrap(~ defClean2$Statistic)
ggplot(data=defClean2, aes(defClean2$Count)) + 
  geom_histogram(binwidth = 1) +
  facet_wrap(~ defClean2$Statistic)


```


Response: All the items which have high frequency in 0 are rare skill. Turnover_FF, Turnover_FR, Turnoner_INT, Pass_QBH are rare skills.


```{r}

defClean_3= subset(defClean2, defClean2$Statistic == "Tackles_Solo")
defClean_ks= subset(defClean_3, defClean_3$Opponent_Opponent =="Kansas State")
defClean_wv= subset(defClean_3, defClean_3$Opponent_Opponent == "West Virginia")
head((defClean_wv))


joined_df <- merge(defClean_ks, defClean_wv, by.x = "Name", 
             by.y = "Name", all.x = TRUE, all.y = TRUE)

joined_df
ggplot(joined_df, aes(x=joined_df$Count.x, y=joined_df$Count.y)) +
  geom_point(size=2, shape=23)

```


Response: Largernumber of solo tackles happened against West Virginia.


```{r}
head(bioClean)

```

## Part 2.2
```{r}

bioClean2 = bioClean %>% separate(Hometown, c("City", "State"), sep = ",")
head(bioClean2)
w = table(bioClean2$State)
t = as.data.frame(w)
t

```


Response: The 'Hometown' column has been separated into two columns, City & State. 
The frequency in the other table shows the number of players from each state. 

## Part 3: Problem 1
```{r}


join_bio_off <- merge(offClean, bioClean, by = "Name"
          )

join_bio_off %>% distinct(Name,Hometown)
```
### Problem 2
```{r}

join_bio_off <- merge(offClean, bioClean, by = "Name"
          )
join_bio_off

plot(join_bio_off$Weight, join_bio_off$Receiving_YDS, main="Scatterplot Example",
   xlab="Weight ", ylab="RY", pch=19)

```
Response: According to the scatterplot, there is almost no relation between the weight of the player and the recieving yards. 


