---
title: "Lab4"
author: "Tyler Smith & Ritika Agrawal"
date: "March 11, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#File Read In
library(readxl)
library(tidyr)
library(dplyr)
library(ggplot2)
defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
str(bio)
```

# Part One  

```{r}
#Part One

#i
defClean <- defense %>%
  mutate(Name = as.factor(Name),Opponent_Opponent=as.factor(Opponent_Opponent)) %>%
  mutate_at(c("Tackles_Solo","Tackles_ASST","Tackles_TFL","Tackles_Sack","Turnover_FF","Turnover_FR","Turnover_INT","Pass_QBH","Pass_PB"),as.numeric)

offClean <- offense %>%
  mutate(Name = as.factor(Name),Opponent_Opponent=as.factor(Opponent_Opponent)) %>%
  replace_na(list(.=0)) %>%
  mutate_at(c("Rushing_ATT","Rushing_YDS","Rushing_TD","Receiving_REC","Receiving_YDS","Receiving_TD","Passing_CMP-ATT","Passing_YDS","Passing_TD","Passing_INT"),as.numeric)

bioClean <- bio %>% 
  separate(col=Height,into=c("Feet","Inches"),sep="-") %>%
  mutate(Name=as.factor(Name),Hometown=as.factor(Hometown),Feet=as.numeric(Feet),Inches=as.numeric(Inches),Weight=as.numeric(Weight)) %>%
  mutate(Height_Inch = Feet*12+Inches) %>%
  select(-Feet,-Inches)

str(defClean)
str(offClean)
str(bioClean)

```
## Part 2
```{r}
defClean2 = pivot_longer(
  defClean,
  cols = Tackles_Solo:Pass_PB,
  names_to = "Statistic",
  values_to = "Count",
  values_drop_na = TRUE
)
head(defClean2)



```

Response: A new column, 'Statistic' has been added. 


```{r}
#qplot(defClean2$Count, geom="histogram") + facet_wrap(~ defClean2$Statistic)
ggplot(data=defClean2, aes(defClean2$Count)) + 
  geom_histogram(binwidth = 1) +
  facet_wrap(~ defClean2$Statistic)


```


Response: All the items which have high frequency in 0 are rare skill. Turnover_FF, Turnover_FR, Turnoner_INT, Pass_QBH are rare skills.


```{r}

defClean_3= subset(defClean2, defClean2$Statistic == "Tackles_Solo")
defClean_ks= subset(defClean_3, defClean_3$Opponent_Opponent =="Kansas State")
defClean_wv= subset(defClean_3, defClean_3$Opponent_Opponent == "West Virginia")
head((defClean_wv))


joined_df <- merge(defClean_ks, defClean_wv, by.x = "Name", 
             by.y = "Name", all.x = TRUE, all.y = TRUE)

joined_df
ggplot(joined_df, aes(x=joined_df$Count.x, y=joined_df$Count.y)) +
  geom_point(size=2, shape=23)

```


Response: Largernumber of solo tackles happened against West Virginia.


```{r}
head(bioClean)

```

##Part 2.2
```{r}

bioClean2 = bioClean %>% separate(Hometown, c("City", "State"), sep = ",")
head(bioClean2)
w = table(bioClean2$State)
t = as.data.frame(w)
t

```


Response: The 'Hometown' column has been separated into two columns, City & State. 
The frequency in the other table shows the number of players from each state. 
    

# Part 3
### Problem 1
```{r}
#Part 3
#Problem 1
bo_join <- offClean %>% left_join(bioClean,by="Name")
home <-  bo_join %>%
  select(Name,Hometown) %>% unique
home
```
Above is the full list of where offensive players are from. There are a variety of cities and states represented.
  
    
###Problem 2
```{r}
#Problem 2
prob2 <- bo_join %>%
  group_by(Name) %>%
  summarize(rec_yds=sum(Receiving_YDS))
head(prob2)
prob2_join <- prob2 %>%
  left_join(bo_join,by="Name") %>%
  select(Name,rec_yds,Weight)
plot(Weight~rec_yds,data=prob2_join,main="Scatterplot of Weight and Rec Yards",xlab="Rec. Yards")
```
From the scatterplot above, we can see that there is little-to-no correlation between weight and receiving yards. Many of the points are clustered towards the bottom left corner of weight and receiving yards, while one of the lightest players actually has the most receiving yards on the team. Meanwhile, there is a near vertical line present betgween 100 and 225 yards, so weight really has no bearing on receiving yards.  
    
        
###Problem 3
```{r}
#Problem 3
#,rush_yds = sum(Rushing_YDS,na.rm=T),rush_td = sum(Rushing_TD,na.rm=T)
isu_stats <- offClean %>%
  group_by(Opponent_Opponent) %>%
  summarize(rush_att = sum(as.numeric(Rushing_ATT),na.rm=T),rush_yds = sum(Rushing_YDS,na.rm=T),rush_td = sum(Rushing_TD,na.rm=T))

head(isu_stats)
isu2 <- offClean %>%
  group_by(Opponent_Opponent)
isu_stats
purdy <- offClean %>%
  filter(Name=="Purdy, Brock") %>%
  select(Opponent_Opponent,Passing_YDS)
combined <- isu_stats %>%
  left_join(purdy,by="Opponent_Opponent")
head(combined)
par(mfrow=c(1,3))
plot(data=combined,rush_att~Passing_YDS,main="How Purdy's Passing yards affects Rush Attempts",ylab="Rushing Attempts",xlab="Purdy's Passing Yards")
plot(data=combined,rush_yds~Passing_YDS,main="How Purdy's Passing yards affects Rush Yards",ylab="Rushing Yards",xlab="Purdy's Passing Yards")
plot(data=combined,rush_td~Passing_YDS,main="How Purdy's Passing yards affects Rush Touchdowns",ylab="Rushing Touchdowns",xlab="Purdy's Passing Yards")

```
  
Above are visuals demonstrating how Brock Purdy's passing yards affect Iowa State's rushing numbers. From the first plot, we see that the three games that Iowa State ran the ball the most, Purdy had less than 300 passing yards, a mark he hit in just under half of the games played. The rushing yards showed similar trends, but still appeared to be slightly more random than rushing attempted. The rushing touchdowns seemed to have little to do with Purdy's passing yards. Purdy's most pass yards on the season also matched the most rushing touchdowns, while two of his bottom-four games in terms of passing yards had four rushing touchdowns.  
    
      
###Problem 4
```{r}
#Problem 4
int_forced <- defClean %>%
  group_by(Opponent_Opponent) %>%
  summarize(def_int = sum(as.numeric(Turnover_INT),na.rm=T))
int_committed <- offClean %>%
  group_by(Opponent_Opponent) %>%
  summarize(off_int = sum(as.numeric(Passing_INT),na.rm=T))
int <- int_forced %>%
  left_join(int_committed,by="Opponent_Opponent")
int
ggplot(int,aes(x=off_int,y=def_int,main="Plot of Interceptions in 2019 Cyclone Football",ylab="Interceptions Forced",xlab="Interceptions Allowed"))+geom_point()+geom_text(aes(label=Opponent_Opponent))


#plot(data=int,def_int~off_int,main="Plot of Interceptions in 2019 Cyclone Football",ylab="Interceptions Forced",xlab="Interceptions Allowed")

```
The question we wanted to answer is:
Is there any correlation between the number of interceptions forced and committed in a game? (Potentially due to weather conditions in a game)  
  
Answer: The plot above reveals that there were three games in 2019 in which both team had at least one interception allowed. In the two games that featured one team having two or more interceptions, the other team also had an interception, which does spark interest as to whether or not there is some correlation.  

Diving deeper into the analysis of the 3+ interception games, the box scores revealed that both games had little-to-no wind and both had temperatures above 50 degrees, which implies weather may not be a huge factor. However, both of these instances happened in Jack Trice Stadium, so perhaps there is something about the stadium that may impact this trend.
  

